---
title: "Stat 382 - HW2"
author: "Max Martinelli"
date: "February 8, 2015"
output: html_document
background: ('#E6F8F3')
---

------------------
# Overview

## Instructions

Write a `SAS` Macro and `R` function to perform the following task.

For a given distribution with two parameters (it can be normal, uniform, or gamma), and an integer $n$:

1. Generate an $n \times n$ symmetric random matrix (say $A$). In this matrix, the elements are random variable from the given distribution. Notice that the matrix must be symmetric.

2. Compute $A^{10}$. You should think about the best way to compute this. Try to save the computation time. Hint: use eigenvalue decomposition. 

<br>



## Preliminaries 

I used some functions from the base package, and made a successful attempt to not use additional libraries such as `Matrix`. While I could have manually built a function to compute the $VDV^{-1}$ decomposition of a matrix, I felt my use of `svd` was appropriate. I initially wrote code to format a symmetric matrix, but the code used and cited below was very elegant and significantly shorter than my approach. 

Surprisingly, the base package of `R` does not have a method to compute the power of a matrix. This means to verify my work would require a bit of work from the user. I do not recommend checking my results with other libraries, especially since other approaches use $VDV^{-1}$ decomposition to compute powers of a matrix (as opposed to traditional matrix multiplication). I have attempted to verify the accuracy of my work in the final section of this page.

```{r}
set.seed(30)
```
<br>

# Approach

## Step 1
First, let us create a function `n_by_n_sym_rand_matrix` that fits the description. 
```{r}

n_by_n_sym_rand_matrix = function(dist, n, para_1, para_2){ 
  
  # This function takes the name of a distribution as a string
  # to compute n-squared random variables given two parameters
  # acceptable inputs are "normal", "uniform", and "gamma"
  
  if (dist == "normal") {
    values = rnorm(n = n**2, mean = para_1, sd = para_2)}
  if (dist == "uniform") {
    values = runif(n = n**2, min = para_1, max = para_2)}
  if (dist == "gamma") {
    values = rgamma(n = n**2, shape = para_1, rate = para_2)}
  
  A = matrix(data = values ,nrow = n, ncol = n)
  A[lower.tri(A)] = t(A)[lower.tri(A)] 
  
  # The above line of code was found at the following link:
  #     http://stackoverflow.com/a/21825946
  # Credit to StackOverflow user: user3318600
  
  return(A)
}

```
<br>

## Step 2

Now, we create a second function `fast_mat_exp` which will raise the matrix to an inputted power.


```{r}

fast_mat_exp = function(mat, exp) {  
  
  # This function takes a matrix and raises it to the exp power
  # This is done using the svd function to obtain the matrix's
  # VDV^-1 decomposition.
  
  decomp = svd(mat)
  D = diag(decomp$d)
  result = decomp$u %*% D**exp %*% solve(decomp$u)
  
  return(result)
}
```

<br>

# Testing

## Instructions 

Test your `SAS` Macro and `R` function with the following set up:

1. Normal distribution with $\mu =1$ and $\sigma^2 = 4$, $n=10$
2. Uniform distribution from $[0, 3]$, and $n=15$
3. Gamma distribution with $\alpha=1$, $\beta=1$, and $n=20$

<br>

## Results

### Test 1

```{r}
A_1 = n_by_n_sym_rand_matrix(dist = "normal", n=10, para_1=1, para_2=4**.5)
round(A_1,2) # preview the rounded A_1 matrix to two decimal places
A_1_to_10 = fast_mat_exp(mat=A_1, exp=10) #construct A^10
# preview the first 5 rows and 5 columns of A^10 with scientific notation
format(A_1_to_10[1:5,1:5], scientific=TRUE, digits = 3) 
```
Now, let us test the accuracy of our `fast_mat_exp` function. We will do this by manually computing $A_1^{10}$ and then comparing the rounded entries to two decimal places. This is necessary to deal with errors that might rise from floating point.
```{r}
A_1_to_5_manual = A_1 %*% A_1 %*% A_1 %*% A_1 %*% A_1
A_1_to_10_manual = A_1_to_5_manual %*% A_1_to_5_manual 
round(A_1_to_10_manual,2) == round(A_1_to_10,2)
```

<br>



### Test 2

```{r}
A_2 = n_by_n_sym_rand_matrix(dist = "uniform", n=15, para_1=0, para_2=3)

# preview the first 10 rows and 10 columns rounded A_2 matrix to two decimal places
round(A_2[1:10,1:10],2) 
A_2_to_10 = fast_mat_exp(mat=A_2, exp=10) #construct A^10
# preview the first 5 rows and 5 columns of A^10 with scientific notation
format(A_2_to_10[1:5,1:5], scientific=TRUE, digits = 3) 
```
Again, we will test the accuracy of our computation. We will do this by manually computing the $A_2^{10}$ and then comparing the rounded entries to one decimal place. Since these matrices are very large, let us count the entries that are equal, and present it as a present it as a table.

```{r}
A_2_to_5_manual = A_2 %*% A_2 %*% A_2 %*% A_2 %*% A_2
A_2_to_10_manual = A_2_to_5_manual %*% A_2_to_5_manual 
results_2 = (round(A_2_to_10_manual,1) == round(A_2_to_10,1))
table(results_2)
```
We see even with rounding to one decimal place, we still have some entries that are off by a bit.

<br>



### Test 3

```{r}
A_3 = n_by_n_sym_rand_matrix(dist = "gamma", n=20, para_1=1, para_2=1)

# preview the first 10 rows and 10 columns rounded A_3 matrix to two decimal places
round(A_3[1:10,1:10],2) 
A_3_to_10 = fast_mat_exp(mat=A_3, exp=10) #construct A^10
# preview the first 5 rows and 5 columns of A^10 with scientific notation
format(A_3_to_10[1:5,1:5], scientific=TRUE, digits = 3) 
```
Again we will test the accuracy of our function to one decimal place.
```{r}
A_3_to_5_manual = A_3 %*% A_3 %*% A_3 %*% A_3 %*% A_3
A_3_to_10_manual = A_3_to_5_manual %*% A_3_to_5_manual 
results_3 = (round(A_3_to_10_manual,1) == round(A_3_to_10,1))
table(results_3)
```

Here we see more discrepancies due to floating point errors. However, the errors are likely to be relatively small.