---
title: "Stat 382 - HW3"
author: "Max Martinelli"
date: "February 25, 2015"
output: html_document
---

------------------

<center> <h2>Overview</h2> </center>

### Instructions

For each problem, write down the hypothesis and also write your conclusion. You should also attach the codes. The level of significance is $0.05$. 

<br>



### Preliminaries 

For this assignment, we will use the [Fisher's Exact Test](http://en.wikipedia.org/wiki/Fisher%27s_exact_test) and the [Chi-square test](http://en.wikipedia.org/wiki/Chi-square_test). We denote the odds ratio as $\theta$. There are six problems. See the document in the [repository](https://github.com/mxmrt/Stat_382/tree/master/hw3) for more details. 


<br>

<center> <h2>Problems</h2> </center>

### Problem 1
#### Problem Restatement

> Is it statistically significant that the proportion of chipmunks trilling is higher when they are closer to their burrow? 

<br>

#### Hypothesis

Our hypotheses are $H_0: \theta = 1$ and $H_1: \theta > 1$. We input the data and use `fisher.test` with the argument `alternative = 'greater'`.

<br>

#### Testing

```{r}

trills=array(c(16, 8, 3, 18),
      dim = c(2, 2),
      dimnames = list(
          trilled = c("did", "did not"),
          distance = c("10 M", "100 M")))

fisher.test(trills, alternative = "greater")

```

<br>

#### Analysis 
Based on the $p-value$ we find the results to be statistically significant and we reject $H_0$.

<br>



### Problem 2
#### Problem Restatement

> Is there no evidence that the two species of birds use the substrates in different proportions?

<br>

#### Hypothesis

Our hypotheses are $H_0: \theta = 1$ and $H_1: \theta \neq 1$. We input the data and use `fisher.test` with the argument `alternative = 'two.sided'`.

<br>

#### Testing

```{r}

substrate=array(c(15, 20, 14, 6, 8, 5, 7, 1),
                dim = c(4, 2),
                dimnames = list(
                    type = c("Vegetation", "Shoreline", "Water", "Structures"),
                    bird = c("Heron", "Egret")))

fisher.test(substrate, alternative = "two.sided")

```

<br>

#### Analysis 
Based on the $p-value$ we do not find the results to be statistically significant and we fail to reject $H_0$.

<br>






### Problem 3
#### Problem Restatement

> Is there a significant difference in synonymous/replacement ratio between polymorphisms and fixed differences?

<br>

#### Hypothesis

Our hypotheses are $H_0: \theta = 1$ and $H_1: \theta \neq 1$. We input the data and use `fisher.test` with the argument `alternative = 'two.sided'`.

<br>

#### Testing

```{r}

gene=array(c(43, 17, 2, 7),
                dim = c(2, 2),
                dimnames = list(
                    fixity = c("polymorphic", "fixed"),
                    synonymicity = c("synonymous", "replacement")))

fisher.test(gene, alternative = "two.sided")

```

<br>

#### Analysis 
Based on the $p-value$ we find the results to be statistically significant and we reject $H_0$.

<br>









### Problem 4
#### Problem Restatement

> We want to check whether being a man or a woman (columns) is independent of having voted in the last election (rows). In other words is "sex and voting independent"?

<br>

#### Hypothesis

Our hypotheses are $H_0:$ the variables are independent and $H_1:$ they are not independent. We input the data and use `chisq.test`.

<br>

#### Testing

```{r}

vote=array(c(2792, 1486, 3591, 2131),
                dim = c(2, 2),
                dimnames = list(
                    position = c("voted", "didn't Vote"),
                    sex = c("Men", "Women")))

chisq.test(vote)

```

<br>

#### Analysis 
Based on the $p-value$ we find the results to be statistically significant and we reject $H_0$.

<br>










### Problem 5
#### Problem Restatement

> What would we conclude (about this die)?

<br>

#### Hypothesis

Our hypotheses are $H_0:$ the variables are independent and $H_1:$ they are not independent. We input the data and use `chisq.test`.

<br>

#### Testing

```{r}

die=array(c(42, 55, 38, 57, 64, 44),
                dim = c(6, 1),
                dimnames = list(
                    face = c("1", "2", "3", "4", "5", "6"),
                    count = c("observed")))

chisq.test(die)

```

<br>

#### Analysis 
Based on the $p-value$ we do not find the results to be statistically significant and we fail to reject $H_0$.

<br>







### Problem 6
#### Problem Restatement

> Based on the data, what is your conclusion (pertaining to 
Calendar Effect for hockey players and date of birth)?

<br>

#### Hypothesis

Our hypotheses are $H_0:$ the variables are independent and $H_1:$ they are not independent. We input the data and use `chisq.test`.

<br>

#### Testing

```{r}

hockey=array(c(84, 77, 35, 34),
                dim = c(4, 1),
                dimnames = list(
                    quarter = c("Jan. to March", "April to June", 
                                "July to Sept.", "Oct. to Dec."),
                    players = c("count")))

chisq.test(hockey)

```

<br>

#### Analysis 
Based on the $p-value$ we find the results to be statistically significant and we reject $H_0$.

<br>



