---
title: "Stat 382: Exam 1, Problem 1"
author: "Max Martinelli"
date: "February 13, 2015"
output: html_document
---
##### Updates:
* *2/13/15, 8:00pm: Changed document title and format. No changes to computation or code were made.* 

### Step 1
First, we fill in a $10 \times 10$ matrix with all results from a gamma random variable:
```{r}

gamma_data = rgamma(n = 10**2, shape = 1, rate = 2)
mat_1 = matrix(data = gamma_data ,nrow = 10, ncol = 10)
mat_1
```

### Step 2
Next, we fill in a $10 \times 10$ matrix with all results from a normal random variable:
```{r}
norm_data = rnorm(n = 10**2, mean = -1, sd = 4**.5)
mat_2 = matrix(data = norm_data ,nrow = 10, ncol = 10)
mat_2
```

### Step 3
Next, we fill in a $10 \times 10$ matrix with all results from a uniform random variable:
```{r}
uni_data = runif(n = 10**2, min = 0, max = 3)
mat_3 = matrix(data = uni_data ,nrow = 10, ncol = 10)
mat_3
```


### Step 4
Finally, we piece together the needed components from each of these three matrices for our final result.

```{r}
result = mat_1
result[lower.tri(result)] = mat_2[lower.tri(mat_2)]
result[upper.tri(result)] = mat_3[upper.tri(mat_3)]
result
```





