---
title: "Stat 382: Exam 1, Problem 2"
author: "Max Martinelli"
date: "February 13, 2015"
output: html_document
---
##### Updates:
* *2/13/15, 8:00pm: Changed document title, and fixed non-computational error that displayed collective BMI mean instead of the mean for the subset of females.* 

### Initial Step
First, we load the data into `R` and preview the data.

```{r}
data = read.csv('/Users/Max/Desktop/Exam1.csv')
head(data)
```

### Part 1
Compute the mean, standard deviation, 5th, 25th, median, 75th, 95th quantile, as well as 95% confidence interval of the mean of BMI.  

```{r}
bmi_mean = mean(data$BMI); bmi_mean
bmi_sd = sd(data$BMI); bmi_sd
bmi_quantile = quantile(data$BMI, probs = c(.05, .25, .5, .75, .95))
bmi_quantile
error <- qt(0.975,df=length(data$BMI)-1)*sd(data$BMI)/sqrt(length(data$BMI))
error
conf_interval = bmi_mean + c(-1,1)*error
conf_interval
```

### Part 2
Compute 90% confidence interval of the proportion of female.

```{r}
fem_data = data[data$FEMALE==1,]
fem_bmi_mean = mean(fem_data$BMI); fem_bmi_mean
fem_error <- qt(0.95,df=length(fem_data$BMI)-1)*sd(fem_data$BMI)/sqrt(length(fem_data$BMI))
fem_error
fem_conf_interval = fem_bmi_mean + c(-1,1)*fem_error
fem_conf_interval
```






